---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


## Overview

The `boilerplate` package provides tools for managing and generating standardised text for methods and results sections of scientific reports. It handles template variable substitution and supports hierarchical organization of text through dot-separated paths.

## Features

- **Text Template Management**: Create, update, and retrieve reusable text templates
- **Variable Substitution**: Replace `{{variable}}` placeholders with actual values
- **Hierarchical Organisation**: Organise text in nested categories using dot notation (e.g., `statistical.longitudinal.lmtp`)
- **Measures Database**: Special handling for research measures with descriptions, items, and metadata
- **Default Content**: Comes with pre-loaded defaults for common methods sections
- **Quarto/R Markdown Integration**: Generate sections for scientific reports


## Installation

You can install the development version of `boilerplate` from GitHub
with:

```{r, eval = FALSE}
# Install the devtools package if you don't have it already
install.packages("devtools")
devtools::install_github("go-bayes/boilerplate")

```



## Basic Usage

## Installation

```r
# Install from GitHub
# install.packages("devtools")
devtools::install_github("go-bayes/boilerplate")
```

## Basic Usage

```r
library(boilerplate)

# Initialize the default databases
boilerplate_init_text()

# Add a new method entry
methods_db <- boilerplate_manage_text(
  category = "methods",
  action = "add",
  name = "sample_selection",
  value = "Participants were selected from {{population}} during {{timeframe}}."
)

# Generate text with variable substitution
methods_text <- boilerplate_generate_text(
  category = "methods",
  sections = c("sample", "sample_selection"),
  global_vars = list(
    population = "university students",
    timeframe = "2020-2021"
  ),
  db = methods_db,
  add_headings = TRUE
)

cat(methods_text)
```

## Managing Measures

The package provides functions for managing measures and generating formatted text about them:
  
  ```r
# View measures database
# First initialize the database (creates/updates the file)
boilerplate_init_measures()

# Then load the database
measures_db <- boilerplate_manage_measures(action = "list")

# Add a measure
# First, load the measures database
measures_db <- boilerplate_manage_measures(action = "list")

# Add the measure with a flat name (no dots)
measures_db <- boilerplate_manage_measures(
  action = "add",
  name = "anxiety_gad7", # Use underscore instead of dot
  measure = list(
    description = "Anxiety was measured using the GAD-7 scale.",
    reference = "spitzer2006",
    waves = "1-3",
    keywords = c("anxiety", "mental health", "gad"),
    items = list(
      "Feeling nervous, anxious, or on edge",
      "Not being able to stop or control worrying",
      "Worrying too much about different things",
      "Trouble relaxing"
    )
  ),
  db = measures_db
)

# Then reference it with the same flat name
exposure_text <- boilerplate_measures_text(
  variable_heading = "Exposure Variable",
  variables = "anxiety_gad7", # Match the name you used above
  db = measures_db,
  heading_level = 3,
  subheading_level = 4,
  print_waves = TRUE
)# First, load the measures database
measures_db <- boilerplate_manage_measures(action = "list")

# dd the measure
measures_db <- boilerplate_manage_measures(
  action = "add",
  name = "anxiety_gad7", # Use underscore instead of dot
  measure = list(
    description = "Anxiety was measured using the GAD-7 scale.",
    reference = "spitzer2006",
    waves = "1-3",
    keywords = c("anxiety", "mental health", "gad"),
    items = list(
      "Feeling nervous, anxious, or on edge",
      "Not being able to stop or control worrying",
      "Worrying too much about different things",
      "Trouble relaxing"
    )
  ),
  db = measures_db
)

# Then reference it with the same flat name
exposure_text <- boilerplate_measures_text(
  variable_heading = "Exposure Variable",
  variables = "anxiety_gad7", # Match the name you used above
  db = measures_db,
  heading_level = 3,
  subheading_level = 4,
  print_waves = TRUE
)

cat(exposure_text)
```

## Text and Measure Management

The core functions for managing text and measures are:
  
  ```r
# Add text to a database
methods_db <- boilerplate_manage_text(
  category = "methods",
  action = "add",
  name = "data_analysis",
  value = "Data were analyzed using {{software}} version {{version}}.",
  db = methods_db
)

# Get text with variables substituted
analysis_text <- boilerplate_manage_text(
  category = "methods",
  action = "get",
  name = "data_analysis",
  template_vars = list(software = "R", version = "4.2.0"),
  db = methods_db
)

# Add a measure to a database
measures_db <- boilerplate_manage_measures(
  action = "add",
  name = "wellbeing",
  measure = list(
    description = "Wellbeing was measured using a 5-item scale.",
    reference = "smith2020",
    items = list("I am satisfied with my life", "My life has meaning"),
    keywords = c("wellbeing", "satisfaction")
  ),
  db = measures_db
)
str(measures_db$wellbeing)
```

## Complete Workflow Example

This example demonstrates combining multiple components to create a complete methods section:
  
```r
# Initialize databases and define study parameters
boilerplate_init_measures()

measures_db <- boilerplate_manage_measures(action = "list")

# Load the measures database
measures_db <- boilerplate_manage_measures(action = "list")

# Add perfectionism measure
measures_db <- boilerplate_manage_measures(
  action = "add",
  name = "perfectionism",
  measure = list(
    name = "perfectionism",
    description = "Perfectionism was measured using a 3-item scale assessing maladaptive perfectionism tendencies.",
    reference = "rice_short_2014",
    waves = "10-current",
    keywords = c("personality", "mental health"),
    items = list(
      "Doing my best never seems to be enough.",
      "My performance rarely measures up to my standards.",
      "I am hardly ever satisfied with my performance."
    )
  ),
  db = measures_db
)

# Save the updated database
boilerplate_manage_measures(
  action = "save",
  db = measures_db
)
# Define parameters
study_params <- list(
  exposure_var = "perfectionism",
  population = "university students",
  timeframe = "2020-2021",
  sampling_method = "convenience"
)

# Generate methods text for participant selection
sample_text <- boilerplate_methods_text(
  sections = c("sample_selection"),
  global_vars = study_params,
  add_headings = TRUE,
  heading_level = "###",
  db = methods_db
)
cat(sample_text)

# Generate measures text for exposure variable
exposure_text <- boilerplate_measures_text(
  variable_heading = "Exposure Variable",
  variables = "perfectionism",
  heading_level = 3,
  subheading_level = 4,
  print_waves = TRUE, 
  db= measures_db
)

cat(exposure_text)


# Generate text for outcome variables by domain
psych_text <- boilerplate_measures_text(
  variable_heading = "Psychological Outcomes",
  variables = c("anxiety.gad7", "depression.phq9"),
  db = measures_db,
  heading_level = 3,
  subheading_level = 4,
  print_waves = TRUE
)
cat(psych_text)

# Generate statistical methods text
stats_text <- boilerplate_methods_text(
  sections = c("statistical.longitudinal.lmtp"),
  global_vars = list(software = "R version 4.2.0"),
  add_headings = TRUE,
  custom_headings = list("statistical.longitudinal.lmtp" = "LMTP"),
  heading_level = "###",
  db = methods_db
)

# Combine all sections into a complete methods section
methods_section <- paste(
  "## Methods\n\n",
  sample_text, "\n\n",
  "### Variables\n\n",
  exposure_text, "\n",
  "### Outcome Variables\n\n",
  psych_text, "\n\n",
  stats_text,
  sep = ""
)
cat(methods_section)
# Save the methods section to a file that can be included in a Quarto document
# writeLines(methods_section, "methods_section.qmd")
```

## License

This package is licensed under the MIT License.

## Code

Go to: <https://github.com/go-bayes/boilerplate>


## DOI

[![DOI](https://zenodo.org/badge/846820825.svg)](https://zenodo.org/doi/10.5281/zenodo.13370825)


