---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# boilerplate

<!-- badges: start -->
<!-- badges: end -->

Certainly! Here's an improved version for clarity and interest:


The `boilerplate` package offers a lightweight and flexible toolkit for managing, accessing, and compiling boilerplate text templates. Designed to streamline the creation and upkeep of template databases, `boilerplate` allows users to efficiently store, retrieve, and update their content. Initially supporting Markdown, the package will soon extend its capabilities to include LaTeX and HTML, making it a versatile solution for a variety of document generation needs.


## Installation

You can install the development version of `boilerplate` from GitHub with:

``` r
# Install the devtools package if you don't have it already
install.packages("devtools")

# Install boilerplate from GitHub
devtools::install_github("go-bayes/boilerplate")
```

## Example

This is a basic example that shows you how to connect to the boilerplate database and create the necessary tables:

```{r example, eval = FALSE}
library(boilerplate)

# connect to the default database (creates it if it doesn't exist)
conn <- boilerplate_connect_db()

# don't forget to disconnect from the database when done
dbDisconnect(conn)
```
