---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# boilerplate

<!-- badges: start -->
<!-- badges: end -->

The goal of `boilerplate` is to provide a comprehensive suite of tools for managing, accessing, and compiling boilerplate text templates. This package is designed to simplify the creation and maintenance of template databases, enabling users to efficiently store, retrieve, update, and compile content into various formats such as Markdown, LaTeX, and HTML.

## Installation

You can install the development version of `boilerplate` from GitHub with:

``` r
# Install the devtools package if you don't have it already
install.packages("devtools")

# Install boilerplate from GitHub
devtools::install_github("yourusername/boilerplate")
```

## Example

This is a basic example that shows you how to connect to the boilerplate database and create the necessary tables:

```{r example, eval = FALSE}
library(boilerplate)

# Connect to the default database (creates it if it doesn't exist)
conn <- boilerplate_connect_db()

# Don't forget to disconnect from the database when done
dbDisconnect(conn)
```
