---
title: "Bibliography Management with boilerplate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bibliography Management with boilerplate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

The boilerplate package provides powerful bibliography management features that integrate seamlessly with your scientific writing workflow. This vignette demonstrates how to set up centralised bibliography management, validate citations, and ensure consistency across your manuscripts.

# Overview: Why Centralised Bibliography Management?

Managing references across multiple manuscripts can be challenging. Common problems include:

- Inconsistent citation formatting
- Missing references in bibliography files
- Duplicated effort updating references across projects
- Version control conflicts with large .bib files

The boilerplate package solves these problems by:

1. **Centralising** your bibliography in one location (e.g., GitHub)
2. **Caching** bibliography files locally for performance
3. **Validating** that all citations exist in your bibliography
4. **Automating** bibliography distribution to project directories

# Setting Up Bibliography Management

## Step 1: Configure Your Bibliography Source

Add bibliography information to your boilerplate database:

```{r setup-bibliography}
library(boilerplate)

# Load your database
# Initialise with default content if needed
temp_bib <- file.path(tempdir(), "bibliography_example")
boilerplate_init(data_path = temp_bib, create_empty = FALSE, create_dirs = TRUE, confirm = FALSE, quiet = TRUE)
db <- boilerplate_import(data_path = temp_bib, quiet = TRUE)

# Add bibliography configuration
# Using the example bibliography included with the package
example_bib <- system.file("extdata", "example_references.bib", package = "boilerplate")
db <- boilerplate_add_bibliography(
  db,
  url = paste0("file://", example_bib),
  local_path = "references.bib",
  validate = TRUE
)

# Save the updated database
boilerplate_save(db, data_path = temp_bib, confirm = FALSE, quiet = TRUE)
```

## Step 2: Download and Cache the Bibliography

The bibliography is automatically cached for performance:

```{r download-bibliography}
# Download/update bibliography (cached for 7 days by default)
bib_file <- boilerplate_update_bibliography(db)

# Force update if needed
bib_file <- boilerplate_update_bibliography(db, force = TRUE)

# Check cache age
boilerplate_update_bibliography(db, force = FALSE)
#> ℹ Using cached bibliography from /path/to/cache/references.bib
#> ⚠ Bibliography cache is 5.2 days old. Consider using force=TRUE to update.
```

## Step 3: Copy Bibliography to Your Project

When working on a manuscript, copy the bibliography to your project directory:

```{r copy-bibliography}
# Copy to current project
boilerplate_copy_bibliography(db, target_dir = ".")

# Copy and update from source first
boilerplate_copy_bibliography(db, target_dir = ".", update_first = TRUE)

# The bibliography is now available as ./references.bib
```

# Validating References

## Check All Citations Exist

Ensure all citations in your boilerplate text exist in the bibliography:

```{r validate-references}
# Validate references across all text categories
validation <- boilerplate_validate_references(db)

# Check specific categories only
validation <- boilerplate_validate_references(
  db, 
  categories = c("methods", "results")
)

# Review validation results
if (!validation$valid) {
  cat("Missing references:\n")
  print(validation$missing)
}

# See all available references
length(validation$available)
#> [1] 1847  # Example: large bibliography

# See which references are actually used
validation$used
#> [1] "@smith2023"     "@jones2024"     "@doe2022meta"
```

## Handle Missing References

When validation finds missing references:

```{r handle-missing}
# Example validation with missing references
validation <- boilerplate_validate_references(db, quiet = TRUE)

if (length(validation$missing) > 0) {
  cat("Please add these references to your bibliography:\n")
  cat(paste0("- ", validation$missing, "\n"))
  
  # Generate BibTeX entries for missing references
  # (This is a manual process - add to your central .bib file)
  for (ref in validation$missing) {
    cat("\n@article{", gsub("@", "", ref), ",\n", sep = "")
    cat("  title = {},\n")
    cat("  author = {},\n")
    cat("  journal = {},\n")
    cat("  year = {},\n")
    cat("}\n")
  }
}
```

# Integration with Document Generation

## Automatic Bibliography Distribution

When generating text, automatically copy the bibliography:

```{r generate-with-bibliography}
# Generate methods text with automatic bibliography copying
methods_text <- boilerplate_generate_text(
  category = "methods",
  sections = c("sample.default", "statistical.default"),
  global_vars = list(n = 1000),
  db = db,
  copy_bibliography = TRUE,
  bibliography_path = "."  # Copy to project root
)

# The bibliography is now available for your Quarto/R Markdown document
```

## Quarto/R Markdown Integration

In your Quarto or R Markdown document:

```{yaml}
---
title: "My Research Paper"
bibliography: references.bib
---
```

```{r document-setup}
# Your document continues with access to all references
# All citations in boilerplate text will be properly resolved
```

# Advanced Workflows

## Project-Specific Bibliography Subsets

For large bibliographies, create project-specific subsets:

```{r subset-bibliography}
# Get citations used in current project
validation <- boilerplate_validate_references(db)
used_refs <- validation$used

# Read full bibliography
bib_lines <- readLines("references.bib")

# Extract entries for used citations
# Note: This is a simplified example - real implementation would need proper BibTeX parsing
# The function extract_bibtex_entries() is not part of boilerplate
# You would need to implement this or use a BibTeX parsing package

# Example approach (pseudo-code):
# library(RefManageR)  # or another BibTeX parsing package
# bib_data <- ReadBib("references.bib")
# project_bib <- bib_data[used_refs]
# WriteBib(project_bib, "project_references.bib")
```

## Multi-Author Collaboration

For collaborative projects with shared boilerplate:

```{r collaboration-workflow}
# 1. Team lead sets up central bibliography
team_db <- boilerplate_import()
team_db <- boilerplate_add_bibliography(
  team_db,
  url = "https://github.com/our-lab/shared-refs/raw/main/lab_references.bib",
  local_path = "lab_references.bib"
)

# 2. Each team member updates their local cache
bib_file <- boilerplate_update_bibliography(team_db, force = TRUE)

# 3. Validate before submission
validation <- boilerplate_validate_references(team_db)
stopifnot(validation$valid)  # Ensure no missing references
```

## Automated Reference Checking

Add to your CI/CD pipeline:

```{r ci-check}
# .github/workflows/check-references.yml
# Run this check on every pull request

# In R script: check_references.R
library(boilerplate)

db <- boilerplate_import()
validation <- boilerplate_validate_references(db, quiet = TRUE)

if (!validation$valid) {
  stop(
    "Missing references found: ", 
    paste(validation$missing, collapse = ", ")
  )
}

message("All references validated successfully!")
```

# Best Practices

## 1. Maintain a Central Bibliography

- Keep your bibliography in version control (e.g., GitHub)
- Use a consistent naming scheme for citation keys
- Regular updates and maintenance
- Consider using BibTeX for key management (free, high functionality)

## 2. Cache Management

```{r cache-management}
# Check cache location
# The cache directory uses R's standard user directory
cache_dir <- tools::R_user_dir("boilerplate", "cache")

# This provides a platform-independent location that complies with CRAN policies:
# - On Unix-like systems (Mac/Linux): ~/.local/share/boilerplate
# - On Windows: Usually in %LOCALAPPDATA%/boilerplate/boilerplate/cache

# View cached files
if (dir.exists(cache_dir)) {
  list.files(cache_dir, pattern = "\\.bib$")
}

# Clear old cache if needed
if (dir.exists(cache_dir)) {
  old_files <- list.files(
    cache_dir, 
    pattern = "\\.bib$", 
    full.names = TRUE
  )
  if (length(old_files) > 0) {
    old_files <- old_files[file.mtime(old_files) < Sys.Date() - 30]
    if (length(old_files) > 0) file.remove(old_files)
  }
}
```


## 3. Citation Key Conventions

Establish consistent citation keys:

```
@article{bulbulia2024methods,     # author + year + keyword
@book{smith2023statistics,        # clear, memorable keys
@inproceedings{jones2024ml,       # avoid special characters
```

# Troubleshooting

## Common Issues and Solutions

### Bibliography Not Downloading

```{r troubleshoot-download}
# Check URL is accessible
url <- db$bibliography$url
con <- url(url)
open(con)
# If this fails, check network/firewall settings
```

### Cache Issues

```{r troubleshoot-cache}
# Force fresh download
bib_file <- boilerplate_update_bibliography(db, force = TRUE)

# Check cache directory permissions
cache_dir <- tools::R_user_dir("boilerplate", "cache")
if (dir.exists(cache_dir)) {
  file.access(cache_dir, mode = 2)  # 0 = success
}
```

### Validation Errors

```{r troubleshoot-validation}
# debug validation issues
validation <- boilerplate_validate_references(db, quiet = FALSE)

# check specific text for citations
text <- db$methods$sample$default
citations <- grep("@[a-zA-Z0-9_:-]+", text, value = TRUE)
print(citations)
```

# Complete Example Workflow

Here's a complete workflow from setup to document generation:

```{r complete-example}
# 1. init setup (run once)
library(boilerplate)

# init new project
temp_complete <- file.path(tempdir(), "complete_example")
boilerplate_init(data_path = temp_complete, create_empty = FALSE, create_dirs = TRUE, confirm = FALSE, quiet = TRUE)

# import database
db <- boilerplate_import(data_path = temp_complete, quiet = TRUE)

# config bibliography
# use example bibliography included with the package
example_bib <- system.file("extdata", "example_references.bib", package = "boilerplate")
db <- boilerplate_add_bibliography(
  db,
  url = paste0("file://", example_bib),
  local_path = "references.bib"
)

# save configuration
boilerplate_save(db, data_path = temp_complete, confirm = FALSE, quiet = TRUE)

# 2. workflow
# update bibliography if needed
boilerplate_update_bibliography(db)

# copy to project
boilerplate_copy_bibliography(db, ".")

# 3. Before submission
# validate all references
validation <- boilerplate_validate_references(db)

if (validation$valid) {
  message("Ready for submission!")
} else {
  warning("Missing references: ", paste(validation$missing, collapse = ", "))
}

# 4. generate final document
final_text <- boilerplate_generate_text(
  category = "methods",
  sections = c("sample.default", "statistical.default"),
  db = db,
  copy_bibliography = TRUE
)
```

# Summary

The boilerplate package's bibliography management features provide:

1. **Centralised management** - One bibliography, many projects
2. **Automatic distribution** - Bibliography copied when needed
3. **Validation** - Ensure all citations are defined
4. **Caching** - Fast local access with periodic updates
5. **Integration** - Works seamlessly with Quarto/R Markdown

By following this workflow, you can maintain consistent, accurate references across all your manuscripts while reducing duplicate effort and potential errors.
