---
title: "Introduction to boilerplate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to boilerplate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(boilerplate)
```

## Overview

The `boilerplate` package provides tools for managing and generating standardised text for methods and results sections of scientific reports. It uses a unified database system with template variable substitution, making it easy to maintain consistent language across projects while allowing for customisation.

## Key Features

- **Unified Database System**: Manage all content types (methods, results, discussion, measures, appendices, templates) through a single interface
- **Template Variable Substitution**: Use `{{variable}}` placeholders for dynamic content
- **Hierarchical Organization**: Organize content using dot notation (e.g., `methods.statistical.longitudinal`)
- **Batch Operations**: Edit multiple entries at once
- **Safety Features**: Confirmation prompts and backup creation
- **JSON Support**: Human-readable JSON format as an alternative to RDS (see `vignette("boilerplate-json-workflow")`)

## Quick Start

### 1. Initialise Databases

First, create and initialise your boilerplate databases:

```{r init, eval=FALSE}
# Initialise all databases with default content
boilerplate_init(
  create_dirs = TRUE,
  create_empty = FALSE,  # Use FALSE to get example content
  confirm = FALSE,
  quiet = FALSE
)
```

### 2. Import Databases

Import your databases to work with them:

```{r import, eval=FALSE}
# Import all databases as a unified structure
unified_db <- boilerplate_import()

# Or import specific categories
methods_db <- boilerplate_import("methods")
measures_db <- boilerplate_import("measures")
```

### 3. Generate Text

Generate text with variable substitution:

```{r generate, eval=FALSE}
# Generate methods text
methods_text <- boilerplate_generate_text(
  category = "methods",
  sections = c("sample", "measures", "statistical_estimator"),
  global_vars = list(
    n_total = 5000,
    exposure_var = "political_conservative",
    outcome_var = "anxiety",
    population = "New Zealand adults"
  ),
  db = unified_db
)

cat(methods_text)
```

### 4. Generate Measure Descriptions

Create formatted descriptions of your measures:

```{r measures, eval=FALSE}
# Generate measure descriptions
measures_text <- boilerplate_generate_measures(
  variable_heading = "Outcome Variables",
  variables = c("anxiety", "depression"),
  db = unified_db,
  sample_items = 3  # Show only first 3 items
)

cat(measures_text)
```

## Working with Databases

### Viewing Database Content

```{r view, eval=FALSE}
# List all paths in a database
all_paths <- boilerplate_list_paths(unified_db)
head(all_paths)

# Get specific content
sample_text <- boilerplate_get_entry(unified_db, "methods.sample.default")
cat(sample_text)
```

### Adding and Updating Content

```{r update, eval=FALSE}
# Add a new entry
unified_db <- boilerplate_add_entry(
  db = unified_db,
  path = "methods.sample.online",
  value = "We recruited {{n_total}} participants through online platforms.",
  category = "methods"
)

# Update existing entry
unified_db <- boilerplate_update_entry(
  db = unified_db,
  path = "methods.sample.default",
  value = "Our sample consisted of {{n_total}} {{population}}.",
  category = "methods"
)

# Save changes
boilerplate_save(unified_db)
```

### Batch Operations

```{r batch, eval=FALSE}
# Update multiple references at once
unified_db <- boilerplate_batch_edit(
  db = unified_db,
  field = "reference",
  new_value = "smith2024",
  target_entries = c("anxiety", "depression", "stress"),
  category = "measures"
)

# Clean text fields
unified_db <- boilerplate_batch_clean(
  db = unified_db,
  field = "description",
  remove_chars = c("@", "[", "]"),
  trim_whitespace = TRUE,
  category = "measures"
)
```

## Template System

The template system uses `{{variable}}` placeholders that are replaced with actual values:

```{r templates, eval=FALSE}
# Create a template
template_text <- "We analysed data from {{n_total}} {{population}} 
  ({{n_female}} female, {{n_male}} male) with a mean age of {{age_mean}} 
  years (SD = {{age_sd}})."

# Use in your database
unified_db <- boilerplate_add_entry(
  db = unified_db,
  path = "methods.participants.demographics",
  value = template_text,
  category = "methods"
)

# Generate with variables
demographics_text <- boilerplate_generate_text(
  category = "methods",
  sections = "participants.demographics",
  global_vars = list(
    n_total = 1000,
    population = "university students",
    n_female = 600,
    n_male = 400,
    age_mean = 21.3,
    age_sd = 3.2
  ),
  db = unified_db
)
```

## Best Practices

1. **Use Version Control**: Export your databases regularly for version control
   ```{r export, eval=FALSE}
   boilerplate_export(
     db = unified_db,
     output_file = "boilerplate_backup_2024.rds"
   )
   ```

2. **Organise Hierarchically**: Use dot notation for clear organisation
   ```{r organise, eval=FALSE}
   # Good organisation
   "methods.measures.psychological.anxiety"
   "methods.measures.psychological.depression"
   "methods.measures.demographic.age"
   ```

3. **Document Your Variables**: Keep a record of template variables
   ```{r document, eval=FALSE}
   # Add a template reference
   unified_db <- boilerplate_add_entry(
     db = unified_db,
     path = "templates.variable_reference",
     value = "Common variables: {{n_total}}, {{exposure_var}}, {{outcome_var}}",
     category = "template"
   )
   ```

4. **Use Meaningful Names**: Choose clear, descriptive names for entries

## Advanced Features

### Measure Standardisation

```{r standardise, eval=FALSE}
# Standardise measure entries
unified_db$measures <- boilerplate_standardise_measures(
  unified_db$measures,
  extract_scale = TRUE,
  clean_descriptions = TRUE
)

# Generate quality report
boilerplate_measures_report(unified_db$measures)
```

### Database Merging

```{r merge, eval=FALSE}
# Merge databases from different sources
merged_db <- boilerplate_merge_databases(
  db1 = project1_db,
  db2 = project2_db,
  db1_name = "Project 1",
  db2_name = "Project 2"
)
```

## Getting Help

For more information and examples, see:

- Package documentation: `help(package = "boilerplate")`
- Function help: `?boilerplate_generate_text`
- GitHub repository: https://github.com/go-bayes/boilerplate
- Example files: `system.file("examples", package = "boilerplate")`
- Example data: `system.file("extdata", package = "boilerplate")`
