---
title: "Managing Multiple Projects with boilerplate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Managing Multiple Projects with boilerplate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(boilerplate)
```

## Introduction

Starting with version 1.1.0, `boilerplate` supports **projects**: independent collections of boilerplate content under one data path. Projects are useful when:

- You want to keep your personal boilerplate separate from shared content
- You're working with colleagues who have their own boilerplate collections
- You manage boilerplate for different research areas or contexts
- You want to experiment without affecting your main boilerplate database

## Project Basics

### Default Project

For backward compatibility, all boilerplate operations use a "default" project if you don't specify one:

```{r eval=FALSE}
# These are equivalent:
boilerplate_init()
boilerplate_init(project = "default")

# Your existing code continues to work unchanged
db <- boilerplate_import()
```

### Creating a New Project

To create a new project, simply specify the project name when initializing:

```{r eval=FALSE}
# Create a project for your lab's shared content
boilerplate_init(
  project = "lab_shared",
  categories = c("methods", "measures"),
  create_dirs = TRUE,
  confirm = FALSE
)

# Create a project for content from a colleague
boilerplate_init(
  project = "smith_measures",
  categories = "measures",
  create_dirs = TRUE,
  confirm = FALSE
)
```

### Working with Projects

All core functions accept a `project` parameter:

```{r eval=FALSE}
# Import from a specific project
lab_db <- boilerplate_import(project = "lab_shared")

# Save to a specific project
boilerplate_save(lab_db, project = "lab_shared")

# Export from a specific project
boilerplate_export(
  db = lab_db,
  project = "lab_shared",
  format = "json"
)
```

## Project Organization

Projects are stored in a structured directory hierarchy:

```
boilerplate/
└── projects/
    ├── default/
    │   └── data/
    │       └── boilerplate_unified.json
    ├── lab_shared/
    │   └── data/
    │       └── boilerplate_unified.json
    └── smith_measures/
        └── data/
            └── boilerplate_unified.json
```

## Cross-Project Operations

### Listing Projects

To see all available projects:

```{r eval=FALSE}
# Simple list
projects <- boilerplate_list_projects()
print(projects)

# With details
boilerplate_list_projects(details = TRUE)
```

### Copying Between Projects

The `boilerplate_copy_from_project()` function enables selective copying of content between projects:

```{r eval=FALSE}
# Copy specific measures from a colleague's project
boilerplate_copy_from_project(
  from_project = "smith_measures",
  to_project = "default",
  paths = c("measures.anxiety", "measures.depression"),
  confirm = FALSE
)

# Copy with a prefix to avoid naming conflicts
boilerplate_copy_from_project(
  from_project = "smith_measures",
  to_project = "default",
  paths = c("measures.anxiety", "measures.depression"),
  prefix = "smith_",  # Results in smith_anxiety, smith_depression
  confirm = FALSE
)
```

### Conflict Handling

When copying between projects, choose one of three conflict strategies:

```{r eval=FALSE}
# Skip conflicting entries (default)
boilerplate_copy_from_project(
  from_project = "lab_shared",
  to_project = "default",
  paths = "methods.sampling",
  merge_strategy = "skip"
)

# Overwrite existing entries
boilerplate_copy_from_project(
  from_project = "lab_shared",
  to_project = "default",
  paths = "methods.sampling",
  merge_strategy = "overwrite",
  confirm = TRUE  # Will ask for confirmation
)

# Rename conflicting entries
boilerplate_copy_from_project(
  from_project = "lab_shared",
  to_project = "default",
  paths = "methods.sampling",
  merge_strategy = "rename"  # Creates sampling_2, sampling_3, etc.
)
```

## Practical Workflow Example

Here's a complete workflow for managing content from multiple sources:

```{r eval=FALSE}
# 1. Initialise your main project (if not already done)
boilerplate_init(
  project = "my_research",
  create_dirs = TRUE,
  confirm = FALSE
)

# 2. Create a project for colleague's content
boilerplate_init(
  project = "colleague_jane",
  categories = "measures",
  create_dirs = TRUE,
  confirm = FALSE
)

# 3. Import and add content to colleague's project
jane_db <- list(
  measures = list(
    work_stress = list(
      description = "Work stress scale from Jane's lab",
      items = 15,
      reference = "Smith2023"
    ),
    burnout = list(
      description = "Burnout inventory",
      items = 22,
      reference = "Smith2022"
    )
  )
)

boilerplate_save(jane_db, project = "colleague_jane")

# 4. Selectively import what you need
boilerplate_copy_from_project(
  from_project = "colleague_jane",
  to_project = "my_research",
  paths = "measures.work_stress",  # Only import the work stress scale
  prefix = "jane_",  # Avoid conflicts
  confirm = FALSE
)

# 5. Verify the import
my_db <- boilerplate_import(project = "my_research")
names(my_db$measures)  # Should include "jane_work_stress"
```

## Best Practices

1. **Use descriptive project names**: Instead of "project1", use "smith_lab_measures" or "meta_analysis_2024"

2. **Document project sources**: Keep notes about where content came from
   ```{r eval=FALSE}
   # Add source information to imported content
   my_db$measures$jane_work_stress$source <- "Jane Smith's lab, imported 2024-01-15"
   boilerplate_save(my_db, project = "my_research")
   ```

3. **Regular backups**: Projects are independent; back them up separately
   ```{r eval=FALSE}
   boilerplate_export(
     db = my_db,
     project = "my_research",
     format = "json",
     output_file = paste0("backup_my_research_", Sys.Date(), ".json")
   )
   ```

4. **Use prefixes for external content**: This prevents naming conflicts and makes sources clear

## Migration from Pre-1.1.0

If you're upgrading from an earlier version, your existing boilerplate data will continue to work:

- Existing data in `boilerplate/data/` is treated as the "default" project
- No changes to your code are required
- You can start using projects whenever you're ready

## Summary

Projects in boilerplate provide a flexible way to:
- Keep different collections of content separate
- Share content with colleagues without mixing databases
- Experiment with new content without affecting your main database
- Maintain clear organisation as your boilerplate collection grows

The project system is entirely optional - if you don't need multiple collections, simply continue using boilerplate as before, and everything will work with the "default" project.
