---
title: "Introduction to boilerplate: Practical Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to boilerplate: Practical Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Why Boilerplate?

Have you ever found yourself:
- Copy-pasting methods sections between papers?
- Struggling to keep consistent language across co-authored manuscripts?
- Manually updating sample sizes in multiple places?
- Losing track of which version of your methods description is the latest?

The `boilerplate` package solves these problems by treating your scientific text as data that can be managed, versioned, and dynamically generated.

# Real-World Example: A Multi-Study Project

Let's walk through a realistic example. Dr. Sarah Chen is conducting a series of studies on workplace well-being. She needs to:

1. Maintain consistent methods descriptions across 3 papers
2. Update sample sizes and demographics dynamically
3. Ensure all co-authors use the same measure descriptions
4. Keep track of citations across all manuscripts

Here's how `boilerplate` helps:

```{r example-setup}
library(boilerplate)

# Sarah's project structure
# Note: In practice, use here::here() for project paths or specify full paths
# Example with here package: project_root <- here::here()
# Example with full path: project_root <- "/Users/sarah/Research/workplace_wellbeing_project"
project_root <- tempfile("workplace_wellbeing_project")  # For this example
dir.create(project_root, recursive = TRUE)
```

## Setting Up the Project

```{r project-setup}
# Initialise boilerplate for the entire project
boilerplate_init(
  data_path = file.path(project_root, "shared/boilerplate"),
  categories = c("methods", "measures", "results", "discussion"),
  create_dirs = TRUE,
  create_empty = FALSE  # Start with helpful examples
)
```

## Creating Reusable Methods Components

Sarah's studies all use similar recruitment methods but different samples:

```{r create-methods}
# Load the database
db <- boilerplate_import(file.path(project_root, "shared/boilerplate"))

# Create a base recruitment method
db <- boilerplate_add_entry(
  db,
  path = "methods.recruitment.online_panel",
  value = paste0(
    "Participants were recruited through {{platform}}, a professional online ",
    "research panel. Eligible participants were {{eligibility_criteria}}. ",
    "The survey took approximately {{duration}} minutes to complete, and ",
    "participants received {{compensation}} as compensation. The final sample ",
    "consisted of {{n}} participants ({{gender_breakdown}})."
  )
)

# Create variants for different platforms
db <- boilerplate_add_entry(
  db,
  path = "methods.recruitment.mturk",
  value = paste0(
    "We recruited {{n}} participants through Amazon Mechanical Turk (MTurk). ",
    "Participation was limited to workers with a HIT approval rate ≥ {{approval_rate}}% ",
    "and ≥ {{min_hits}} completed HITs. Workers received ${{payment}} USD for ",
    "completing the {{duration}}-minute survey. After excluding {{n_excluded}} ",
    "participants who failed attention checks, the final sample included {{n_final}} ",
    "participants (Mage = {{m_age}}, SD = {{sd_age}}; {{pct_female}}% female)."
  )
)

# Save the database
boilerplate_save(db, file.path(project_root, "shared/boilerplate"))
```

## Study 1: Pilot Study

For her pilot study, Sarah generates the methods section:

```{r study1-methods}
# Study 1 parameters
study1_params <- list(
  platform = "Prolific",
  eligibility_criteria = "currently employed full-time and based in the UK",
  duration = 15,
  compensation = "£2.50",
  n = 150,
  gender_breakdown = "52% female, 47% male, 1% non-binary"
)

# Generate methods for Study 1
study1_methods <- boilerplate_generate_text(
  category = "methods",
  sections = c(
    "recruitment.online_panel",
    "design.cross_sectional",
    "analysis.regression"
  ),
  global_vars = study1_params,
  db = db
)

# Save to Study 1 manuscript
cat(study1_methods, file = file.path(project_root, "study1/methods.txt"))
```

## Study 2: Main Study

Six months later, Sarah runs her main study with a larger sample:

```{r study2-methods}
# Study 2 parameters - MTurk this time
study2_params <- list(
  n = 500,
  approval_rate = 95,
  min_hits = 100,
  payment = 4.00,
  duration = 20,
  n_excluded = 47,
  n_final = 453,
  m_age = 34.7,
  sd_age = 10.2,
  pct_female = 58.3
)

# Generate methods for Study 2
study2_methods <- boilerplate_generate_text(
  category = "methods",
  sections = c(
    "recruitment.mturk",
    "design.cross_sectional",
    "measures.job_satisfaction",
    "measures.burnout",
    "analysis.sem"
  ),
  global_vars = study2_params,
  db = db
)
```

# Managing Measures Across Studies

## Creating a Measure Library

Sarah's lab uses standard measures across projects:

```{r measures-library}
# Add burnout measure used across studies
db <- boilerplate_add_entry(
  db,
  path = "measures.burnout.mbi",
  value = list(
    name = "Maslach Burnout Inventory",
    abbreviation = "MBI",
    description = paste0(
      "Burnout was assessed using the {{version}} ({{citation}}). ",
      "This scale consists of {{n_items}} items measuring three dimensions: ",
      "emotional exhaustion ({{n_ee}} items, e.g., '{{example_ee}}'), ",
      "depersonalization ({{n_dp}} items, e.g., '{{example_dp}}'), and ",
      "personal accomplishment ({{n_pa}} items, e.g., '{{example_pa}}'). ",
      "Items are rated on a {{scale}} frequency scale. {{scoring_note}}"
    ),
    items = list(
      emotional_exhaustion = c(
        "I feel emotionally drained from my work",
        "I feel used up at the end of the workday",
        "I feel fatigued when I get up in the morning"
      ),
      depersonalization = c(
        "I feel I treat some recipients as if they were impersonal objects",
        "I've become more callous toward people since I took this job"
      ),
      personal_accomplishment = c(
        "I can easily understand how my recipients feel",
        "I deal very effectively with the problems of my recipients"
      )
    ),
    psychometrics = list(
      reliability = "Internal consistency in the current sample was excellent (α = {{alpha}})",
      validity = "The three-factor structure was confirmed using CFA, χ²({{df}}) = {{chi2}}, CFI = {{cfi}}, RMSEA = {{rmsea}}"
    )
  )
)
```

## Generating Consistent Measure Descriptions

Now any team member can generate consistent measure descriptions:

```{r generate-measures}
# Parameters for MBI
mbi_params <- list(
  version = "Maslach Burnout Inventory - General Survey",
  citation = "Maslach et al., 1996",
  n_items = 16,
  n_ee = 5,
  n_dp = 5,
  n_pa = 6,
  example_ee = "I feel emotionally drained from my work",
  example_dp = "I've become more callous toward people",
  example_pa = "I feel I'm positively influencing people's lives",
  scale = "7-point",
  scoring_note = "Higher scores indicate greater burnout for EE and DP; scores are reversed for PA.",
  alpha = ".91",
  df = 101,
  chi2 = "247.3",
  cfi = ".94",
  rmsea = ".068"
)

# Generate measure description
mbi_description <- boilerplate_generate_text(
  category = "measures",
  sections = "burnout.mbi",
  global_vars = mbi_params,
  db = db
)
```

# Collaborative Features

## Sharing Across Team Members

Sarah shares the boilerplate database with her team:

```{r team-sharing}
# Export to JSON for version control
boilerplate_export(
  db,
  data_path = file.path(project_root, "shared/boilerplate"),
  output_file = "lab_boilerplate_v2.json",
  format = "json"
)

# Team member imports
team_db <- boilerplate_import(
  file.path(project_root, "shared/boilerplate/lab_boilerplate_v2.json")
)
```

## Batch Updates

When reviewer 2 asks for more detail on exclusion criteria:

```{r batch-updates}
# Add exclusion criteria to all recruitment methods
db <- boilerplate_batch_edit(
  db,
  field = "exclusion_note",
  new_value = paste0(
    " Participants were excluded if **before randomisation** they: (a) failed more than {{n_attention}} ",
    "attention check items, (b) completed the survey in less than {{min_time}} ",
    "of the median completion time, or (c) provided nonsensical open-ended responses."
  ),
  target_entries = "methods.recruitment.*",
  category = "methods"
)
```

# Advanced Workflows

## Dynamic Report Generation

Create a function to generate complete methods sections:

```{r dynamic-generation}
generate_study_methods <- function(study_config, db) {
  # Extract sections to include based on study type
  sections <- c(
    paste0("recruitment.", study_config$recruitment_method),
    paste0("design.", study_config$design),
    study_config$measures,
    paste0("analysis.", study_config$analysis)
  )
  
  # Add standard sections
  if (study_config$has_missing_data) {
    sections <- c(sections, "missing.multiple_imputation")
  }
  
  if (study_config$has_power_analysis) {
    sections <- c(sections, "power.post_hoc")
  }
  
  # Generate text
  boilerplate_generate_text(
    category = "methods",
    sections = sections,
    global_vars = study_config$parameters,
    db = db
  )
}

# Use for any study
study3_config <- list(
  recruitment_method = "mturk",
  design = "experimental",
  measures = c("measures.burnout.mbi", "measures.engagement.uwes"),
  analysis = "anova",
  has_missing_data = TRUE,
  has_power_analysis = TRUE,
  parameters = list(
    n = 300,
    approval_rate = 98,
    # ... other parameters
  )
)

study3_methods <- generate_study_methods(study3_config, db)
```

## Quality Control

Ensure consistency across manuscripts:

```{r quality-control}
# Check all papers use current measure descriptions
papers <- c("study1", "study2", "study3")

for (paper in papers) {
  paper_db <- boilerplate_import(
    file.path(project_root, paper, "boilerplate")
  )
  
  # Compare measure descriptions
  differences <- compare_entries(
    paper_db$measures$burnout,
    db$measures$burnout
  )
  
  if (length(differences) > 0) {
    message("Update needed for ", paper, ": ", differences)
  }
}
```

# Integration with R Markdown/Quarto

## Seamless Document Generation

In your Quarto document (`manuscript.qmd`):

````markdown
---
title: "Workplace Well-being Study 3"
author: "Sarah Chen et al."
format: pdf
bibliography: references.bib
---

```{r setup, include=FALSE}
library(boilerplate)
db <- boilerplate_import("shared/boilerplate")

# Study-specific parameters
params <- list(
  n = 300,
  m_age = 36.2,
  sd_age = 11.4,
  pct_female = 54.7,
  # ... more parameters
)
```

# Method

```{r methods, echo=FALSE, results='asis'}
methods <- boilerplate_generate_text(
  category = "methods",
  sections = c(
    "recruitment.mturk",
    "design.experimental",
    "measures.burnout.mbi",
    "measures.engagement.uwes", 
    "analysis.moderation"
  ),
  global_vars = params,
  db = db,
  copy_bibliography = TRUE  # Ensures references.bib is available
)

cat(methods)
```

# Results

```{r results, echo=FALSE, results='asis'}
results_intro <- boilerplate_generate_text(
  category = "results",
  sections = "preliminary.descriptives",
  global_vars = params,
  db = db
)

cat(results_intro)
```
````

# Best Practices

## 1. Organize Your Content Hierarchically

```{r organization}
# Good: Clear hierarchy
"methods.recruitment.online.mturk"
"methods.recruitment.online.prolific"
"methods.recruitment.in_person.lab"
"methods.recruitment.in_person.field"

# Less optimal: Flat structure
"methods.mturk_recruitment"
"methods.prolific_recruitment"
"methods.lab_recruitment"
```

## 2. Use Meaningful Variable Names

```{r variable-naming}
# Good: Self-documenting variables
"{{n_participants}}"
"{{mean_age}}"
"{{pct_female}}"
"{{cronbach_alpha}}"

# Less clear
"{{n}}"
"{{m1}}"
"{{p1}}"
"{{a}}"
```

## 3. Version Control Your Boilerplate

```{r version-control}
# Export with timestamp for versioning
timestamp <- format(Sys.time(), "%Y%m%d")
filename <- paste0("boilerplate_backup_", timestamp, ".json")

boilerplate_export(
  db,
  output_file = filename,
  format = "json"
)

# Track in git
# git add boilerplate_backup_20240115.json
# git commit -m "Boilerplate snapshot after reviewer revisions"
```

## 4. Create Project-Specific Templates

```{r project-templates}
# Create a template for your specific journal's requirements
db <- boilerplate_add_entry(
  db,
  path = "templates.journal.plos_one",
  value = list(
    word_limit = 3500,
    abstract_limit = 300,
    methods_requirements = paste0(
      "PLOS ONE requires detailed methods including: ",
      "{{sampling_procedure}}, {{ethics_statement}}, ",
      "{{data_availability}}, and {{statistical_approach}}."
    )
  )
)
```

# Summary

The `boilerplate` package transforms scientific writing from a copy-paste exercise into a structured, maintainable process. Key benefits:

- **Consistency**: Same descriptions across all papers
- **Efficiency**: Update once, use everywhere
- **Collaboration**: Share standardised text with your team
- **Reproducibility**: Version control your methods
- **Flexibility**: Customize for each study while maintaining standards

Ready to get started? Check out:
- [Getting Started Tutorial](boilerplate-getting-started.html) for a complete walkthrough
- [Bibliography Management](boilerplate-bibliography-workflow.html) for citation handling
- [Measures Workflow](boilerplate-measures-workflow.html) for detailed measure documentation

Your future self (and your collaborators) will thank you!
