---
title: "Version Management and Backup Recovery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Version Management and Backup Recovery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r load-package}
library(boilerplate)
```

## Overview

The boilerplate package provides comprehensive version management features to help you track changes, manage different versions of your databases, and recover from backups when needed.

## Managing Multiple Versions

When working on a project over time, you'll accumulate different versions of your databases. The boilerplate package helps manage these:

```{r list-versions}
# See all available database files
# Using a temporary directory for this example
temp_version <- file.path(tempdir(), "version_example")
boilerplate_init(data_path = temp_version, create_empty = FALSE, create_dirs = TRUE, confirm = FALSE, quiet = TRUE)

files <- boilerplate_list_files(data_path = temp_version)
print(files)

# Check what versions exist for methods
methods_files <- boilerplate_list_files(data_path = temp_version, category = "methods")
```

The `boilerplate_list_files()` function organises files into:
- **Standard files**: Current working versions
- **Timestamped versions**: Saved with timestamps
- **Backup files**: Automatic backups created before overwrites

## Working with Timestamped Versions

Save important milestones with timestamps:

```{r save-with-timestamp}
# Load your database
db <- boilerplate_import(data_path = temp_version, quiet = TRUE)

# Save with timestamp before major revision
boilerplate_save(
  db,
  data_path = temp_version,
  timestamp = TRUE,
  confirm = FALSE,
  quiet = FALSE
)
#> ✔ Saved unified database to: boilerplate_unified_20240115_143022.json

# Later, import specific version
# Note: Replace with your actual timestamped filename
db_milestone <- boilerplate_import(
  data_path = "path/to/your/boilerplate_unified_20240115_143022.json"
)
```

## Backup and Recovery

If something goes wrong, recover from automatic backups:

```{r backup-recovery}
# List available backups
files <- boilerplate_list_files(data_path = temp_version)
# Look at files$backups for backup files

# View latest backup without restoring
backup_db <- boilerplate_restore_backup(data_path = temp_version, category = "methods")

# Restore latest backup as current version
boilerplate_restore_backup(
  data_path = temp_version,
  category = "methods",
  restore = TRUE,
  confirm = TRUE
)
#> ✔ Restored backup from 20240110_120000

# Restore specific backup by timestamp
boilerplate_restore_backup(
  data_path = temp_version,
  category = "methods",
  backup_version = "20240110_120000",
  restore = TRUE,
  confirm = FALSE
)
```

## Version Management Workflow

Here's a typical workflow for managing versions:

```{r workflow-example}
# 1. Check current versions
files <- boilerplate_list_files(data_path = temp_version, category = "methods")

# 2. Save current work with timestamp
methods_db <- boilerplate_import("methods", data_path = temp_version, quiet = TRUE)
boilerplate_save(
  methods_db,
  data_path = temp_version,
  category = "methods",
  timestamp = TRUE,
  confirm = FALSE,
  quiet = TRUE
)

# 3. Make changes
methods_db$new_method <- "New methodology text"
boilerplate_save(methods_db, data_path = temp_version, category = "methods", confirm = FALSE, quiet = TRUE)

# 4. If changes were problematic, restore from backup
boilerplate_restore_backup(data_path = temp_version, category = "methods", restore = TRUE, confirm = FALSE)

# 5. Compare versions
current <- boilerplate_import("methods", data_path = temp_version, quiet = TRUE)
old_version <- boilerplate_import(
  data_path = "path/to/your/methods_db_20240110_120000.json"
)
```

## Best Practices

1. **Regular timestamped saves**: Save with timestamps at key milestones
2. **Document changes**: Use meaningful commit messages when saving
3. **Clean up old versions**: Periodically review and archive old versions
4. **Test before overwriting**: Import and check old versions before replacing

## Integration with Version Control

While the package provides its own versioning, it works well with Git:

```{r git-integration}
# Save timestamped version for Git commit
boilerplate_save(db, data_path = temp_version, timestamp = TRUE, confirm = FALSE, quiet = TRUE)

# Add to Git
# git add boilerplate/data/boilerplate_unified_*.json
# git commit -m "Snapshot before major refactoring"
```

## Troubleshooting

### Finding Lost Work

If you've lost work, check these locations:

```{r troubleshooting}
# List all files including backups
all_files <- boilerplate_list_files(data_path = temp_version)

# Check modification times
# The most recently modified files appear first
print(all_files$timestamped[1:5, c("file", "modified")])
print(all_files$backups[1:5, c("file", "modified")])
```

### Recovering from Corruption

If a database file becomes corrupted:

```{r recovery}
# Try loading backup
backup_db <- tryCatch(
  boilerplate_restore_backup(data_path = temp_version, category = "methods"),
  error = function(e) {
    message("Backup corrupted, trying older version...")
    # List files and manually load an older one
    files <- boilerplate_list_files(data_path = temp_version, category = "methods")
    if (nrow(files$timestamped) > 0) {
      boilerplate_import(data_path = files$timestamped$path[2])
    }
  }
)
```

## Summary

Version management in boilerplate provides:
- Automatic backups before overwrites
- Timestamped saves for milestones
- Easy listing and organisation of versions
- Simple restoration from backups
- Direct import of any version by path

This ensures you never lose work and can always recover from mistakes or explore different versions of your content.