---
title: "Using boilerplate in Quarto: A Minimal Example"
author: "Your Name"
format: 
  html:
    toc: true
execute:
  echo: true
  warning: false
---

## Introduction

This document demonstrates how to use the `boilerplate` package in a Quarto document to manage reusable text templates for scientific writing.

## Setup

First, we'll load the package and initialise our databases in a project-specific directory:

```{r setup}
library(boilerplate)

# Use a project-specific directory to avoid confusion
data_path <- file.path(tempdir(), "boilerplate-example")

# Initialise with clean setup (no backups for this example)
boilerplate_init(
  data_path = data_path,
  create_dirs = TRUE,
  quiet = TRUE
)
```

## Adding Standard Methods Text

Let's add some reusable methods text to our database:

```{r add-methods}
# Add participant recruitment text
boilerplate_add_entry(
  db_path = file.path(data_path, "boilerplate_methods.rds"),
  category = "methods",
  path = "participants.recruitment",
  value = list(
    description = "Standard participant recruitment text",
    text = "We recruited {{n_participants}} participants from {{source}} between {{start_date}} and {{end_date}}. Participants were eligible if they {{eligibility}}. The final sample consisted of {{n_final}} participants ({{pct_female}}% female, mean age = {{mean_age}} years, SD = {{sd_age}})."
  ),
  confirm = FALSE,
  quiet = TRUE
)

# Add data analysis text
boilerplate_add_entry(
  db_path = file.path(data_path, "boilerplate_methods.rds"),
  category = "methods",
  path = "analysis.primary",
  value = list(
    description = "Primary analysis approach",
    text = "We conducted {{analysis_type}} using {{software}}. {{model_description}}. We set alpha at {{alpha}} and used {{correction}} for multiple comparisons."
  ),
  confirm = FALSE,
  quiet = TRUE
)
```

## Adding Measures

Now let's add information about our measures:

```{r add-measures}
# Add a depression measure
boilerplate_add_entry(
  db_path = file.path(data_path, "boilerplate_measures.rds"),
  category = "measures",
  path = "phq9",
  value = list(
    name = "PHQ-9",
    description = "Patient Health Questionnaire-9 for depression screening",
    items = list(
      "Little interest or pleasure in doing things",
      "Feeling down, depressed, or hopeless",
      "Trouble falling or staying asleep, or sleeping too much",
      "Feeling tired or having little energy",
      "Poor appetite or overeating",
      "Feeling bad about yourself",
      "Trouble concentrating on things",
      "Moving or speaking slowly/being fidgety or restless",
      "Thoughts that you would be better off dead"
    ),
    reference = "Kroenke et al., 2001",
    scoring = "Sum of all items (0-27)",
    reliability = "α = .86-.89"
  ),
  confirm = FALSE,
  quiet = TRUE
)
```

## Generating Methods Section

Now we can generate our methods section with specific values:

```{r generate-methods}
# Import the database
db <- boilerplate_import(data_path = data_path, quiet = TRUE)

# Generate participant section
participant_text <- boilerplate_generate_text(
  db_path = file.path(data_path, "boilerplate_methods.rds"),
  category = "methods",
  sections = "participants.recruitment",
  global_vars = list(
    n_participants = 250,
    source = "university students and community members",
    start_date = "January 2024",
    end_date = "March 2024",
    eligibility = "were 18 years or older and fluent in English",
    n_final = 237,
    pct_female = 58,
    mean_age = 32.4,
    sd_age = 11.2
  ),
  quiet = TRUE
)

# Generate analysis section
analysis_text <- boilerplate_generate_text(
  db_path = file.path(data_path, "boilerplate_methods.rds"),
  category = "methods",
  sections = "analysis.primary",
  global_vars = list(
    analysis_type = "multilevel regression analyses",
    software = "R version 4.3.2 with the lme4 package",
    model_description = "We included random intercepts for participants and fixed effects for time, condition, and their interaction",
    alpha = ".05",
    correction = "Bonferroni correction"
  ),
  quiet = TRUE
)
```

### Methods

#### Participants

`r participant_text`

#### Data Analysis

`r analysis_text`

## Generating Measures Section

```{r generate-measures}
measures_text <- boilerplate_generate_measures(
  variable_heading = "Depression Measure",
  variables = "phq9",
  db = db$measures,
  print_keywords = FALSE,
  quiet = TRUE
)
```

### Measures

`r measures_text`

## Batch Operations Example

Let's say we want to update all our methods to reference a new version of R:

```{r batch-update}
# First, let's see what would change (preview mode)
boilerplate_batch_edit(
  db = db$methods,
  field = "text",
  new_value = function(old_text) {
    if (is.character(old_text)) {
      gsub("R version 4.3.2", "R version 4.4.0", old_text)
    } else {
      old_text
    }
  },
  target_entries = "*",
  category = "methods",
  preview = TRUE,
  quiet = TRUE
)
```

## Saving Your Work

When you're done, you can save your unified database:

```{r save-db}
# Save the entire database
boilerplate_save(
  db,
  data_path = data_path,
  confirm = FALSE,
  quiet = TRUE,
  create_backup = FALSE  # No backups for this example
)
```

## Bibliography Integration

The boilerplate package can manage your bibliography file:

```{r bibliography-example}
# Add bibliography to database
db <- boilerplate_add_bibliography(
  db,
  url = "https://raw.githubusercontent.com/go-bayes/templates/main/bib/references.bib",
  local_path = "references.bib"
)

# Generate text with bibliography
methods_with_bib <- boilerplate_generate_text(
  category = "methods",
  sections = "analysis.primary",
  global_vars = list(
    analysis_type = "regression analyses following @smith2023"
  ),
  db = db,
  copy_bibliography = TRUE,
  bibliography_path = ".",
  quiet = TRUE
)

# Validate citations
validation <- boilerplate_validate_references(db, quiet = TRUE)
if (!validation$valid) {
  warning("Missing references found!")
}
```

## Tips for Using boilerplate with Quarto

1. **Project Organization**: Keep your boilerplate data in a project-specific directory (e.g., `.boilerplate-data/`) rather than the default location.

2. **Version Control**: Add your boilerplate database files to git so collaborators can use the same templates.

3. **Bibliography Management**: Use the bibliography features to ensure consistent citations across documents.

4. **Parameters**: Use Quarto parameters for values that change between documents:
   ```yaml
   params:
     n_participants: 250
     study_name: "Study 1"
   ```

4. **Child Documents**: Create separate `.qmd` files for methods, results, etc., and include them as child documents:
   ```markdown
   {{< include _methods.qmd >}}
   ```

5. **Functions**: Write wrapper functions for commonly used text generation:
   ```r
   generate_my_methods <- function(n, start_date, end_date) {
     boilerplate_generate_text(
       db_path = ".boilerplate-data/methods.rds",
       category = "methods",
       sections = c("participants", "procedures", "analysis"),
       global_vars = list(
         n_participants = n,
         start_date = start_date,
         end_date = end_date
       )
     )
   }
   ```

## Conclusion

The `boilerplate` package streamlines scientific writing in Quarto by:

- Centralizing reusable text templates
- Enabling variable substitution for customization
- Supporting hierarchical organization of content
- Providing batch operations for maintenance
- Integrating seamlessly with Quarto's workflow

This approach ensures consistency across documents while maintaining flexibility for project-specific details.
